---
title: 'Dose response workshop: Analysis'
author: "Wenjuan Ma"
date: ''
output:
  pdf_document:
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
header-includes:
- \usepackage{fancyhdr}
- \usepackage[yyyymmdd,hhmmss]{datetime}
- \usepackage{lastpage}
- \usepackage{fontspec}
- \pagestyle{fancy}
- \lhead{Data Analysis for liver drug project}
- \rhead{\today\ \currenttime}
- \cfoot{ }
- \fancyfoot[RE,RO]{\thepage\ of \pageref*{LastPage}}
- \renewcommand{\headrulewidth}{0.4pt}
- \renewcommand{\footrulewidth}{0.4pt}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \lfoot{\texttt{\small \detokenize{`r paste0(getwd(), "/", sub(".Rmd", ".pdf", knitr:::current_input()))`}}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

********************************************************************************
# Document Information
Folder: *`r getwd()`*  

* Script: *`r knitr:::current_input()`*
* Output: *`r sub(".Rmd", ".pdf", knitr:::current_input())`*


. Sponsor: National Institute of Health: 
Michigan State University, East Lansing, MI. 

## Purpose
 

## Revision History
Dates in this section use the MM/DD/YY format. 

* 05/13/2019
* 07/26/2019

## Assumptions
* We use R Markdown to enhance reproducibility. Knitting the R Markdown script
  generates the PDF file containing explanatory text, R code, plus R output 
  (text and graphics) noted in the Document Information section above.
* We will use RStudio (version 1.1.447 or later, www.rstudio.org) to knit this 
  script. The software chain looks like this **Rmd file -> RStudio -> R -> knitr 
  -> pandoc -> MiKTeX -> PDF file**.
* We are using MiKTeX 2.9 (https://miktex.org). 

********************************************************************************
# Setup Tasks

Set global R chunk options (local chunk options will over-ride global options). 

``` {r modified-global_options}
# Global chunk options (must be a stand-alone initial R chunk). 
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = TRUE, warning = TRUE)

# Declare path to this script relative to the project root directory.
here::i_am(
  path = "Dose_Response_App/WORKSHOP_03152022.Rmd"
  )
```



Start from a known state by cleaning the workspace. Then, start timing the 
script and note the default working directory (where this file is located in our 
folder structure). Rmarkdown does not preserve setting the working directory 
across chunks (every chunk defaults to folder where the script is stored).

```{r setup_tasks}
# Clean the workspace (delete all objects in memory & collect garbage).
rm(list = ls())
gc()
options(scipen = 999)
# Start timing the analysis.
print(Timing <- data.frame(Start = Sys.time(), End = NA, Duration = NA))

# Show the default working directory for R chunks in this script.
getwd()
```

## Load Packages & Set Options
Load contributed R packages that we need to get additional functions. 

```{r load_packages}
library(prettyR)          # for stretch_df()
library(descr)            # For freq() crosstab().
options(descr.plot=FALSE) # Make freq() & crosstab() skip plots by default.
library(lubridate)        # For date conversion, eg. ymd(), time_length().
library(sjlabelled)       # For set_label(), get_label()
library(haven)            # For read_sav(), write_sav(), as_factor().
options(tibble.print_max = Inf)  # Show all rows by default. 
options(tibble.width = Inf)      # Show all columns by default
library(plyr)             # For ddply()
library(knitr)            # for kable()
library(car)              # For recode()
library(dplyr)            # For recode(), arrange(), distinct()
library(Hmisc)            # For Cs()
library(foreign)          # For write.foreign()
library(stringr)  # For the str_replace_all() and str_count() functions.
library(tidyr)            # For gather()
library(ggplot2)          # For ggplot()
library(knitr)            # For kable() 
library(kableExtra)       # For kable title
library(stringr)          # For gsub()
library(lattice)
library(ggplot2)
library(reshape2)
library(drc)
library(npreg)
```

## Read in data sete 

```{modified-R}
Data = read.csv("C:/Users/mawenjua/OneDrive - Michigan State University/Workshops/Dose-response curve workshop/data.csv")
```

# Run ANOVA and Extract P-values

```{R}
Data$Drug = as.character(Data$Drug)
Data$Drug = tolower(Data$Drug)
DRUG.NAME = data.frame(unique(Data$Drug))
Drugs_cov = DRUG.NAME
names(Drugs_cov) = c("DRUG.NAME")
Drugs_cov$DRUG.NAME = as.character(Drugs_cov$DRUG.NAME)

for (name in Drugs_cov$DRUG.NAME){
     data = subset(Data, Drug == name)
	   print(summary(aov(lm(data$Response ~ data$CONCEN.uM))))
	   Drugs_cov$pvalue[Drugs_cov$DRUG.NAME == name] = unlist(summary(aov(lm(data$Response ~ data$CONCEN.uM))))[9]} 

for (name in Drugs_cov$DRUG.NAME){
      Drugs_cov$pvaluecut[Drugs_cov$DRUG.NAME == name] = ifelse(Drugs_cov$pvalue[Drugs_cov$DRUG.NAME == name] <= 0.01, "p <= 0.01", "p > 0.01")}
```

# Error bar figure

```{R}
data = subset(Data, Drug == "drug2")
dataagg = cbind(aggregate(data$Response, by = list(data$CONCEN.uM), FUN = mean, na.rm = T), aggregate(data$Response, by = list(data$CONCEN.uM), FUN = sd, na.rm = T)$x)
names(dataagg) = c("CONCEN.uM", "MEAN", "SD")
figure = ggplot(dataagg, aes(x = CONCEN.uM, y = MEAN)) + scale_x_continuous(name = "CONCEN.uM", trans='log')
figure = figure + geom_errorbar(aes(ymin = MEAN - SD, ymax = MEAN + SD), width = .1, color = "green")
figure = figure + geom_line(color = "green") + geom_point(color = "green")
figure = figure + theme_bw() + theme_classic()
```

# Run Does-response curve models for the drugs have significant ANOVA

```{R}
# 3-parameter log-logistic model
# Drug 2
name = "drug2"
data = subset(Data, Drug == "drug2")

m1_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = LL.3u(upper = 100))
summary(m1_d1)
AIC(m1_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = LL.3u(upper = 100)), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]


# 4-parameter log-logistic model
m2_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = LL.4())
summary(m2_d1)
AIC(m2_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = LL.4()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 3-parameter asymptotic regression model
m3_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = AR.3())
summary(m3_d1)
AIC(m3_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = AR.3()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 3-parameter exponential model
m4_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = EXD.3())
summary(m4_d1)
AIC(m4_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = EXD.3()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# two 3 parameter Weibull models
m5_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W1.3())
summary(m5_d1)
AIC(m5_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W1.3()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# two 3-parameter Weibull models
m6_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W2.3())
summary(m6_d1)
AIC(m6_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W2.3()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 4-parameter Weibull models
m7_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W1.4())
summary(m7_d1)
AIC(m7_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W1.4()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 4-parameter Weibull model
m8_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W2.4())
summary(m8_d1)
AIC(m8_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = W2.4()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 3-parameter Gompertz model
m9_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = G.3())
summary(m9_d1)
AIC(m9_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = G.3()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

# 4-parameter Gompertz model
m10_d1 = drm(Response ~ CONCEN.uM, data = na.omit(data), fct = G.4())
summary(m10_d1)
AIC(m10_d1)

ED(drm(Response ~ CONCEN.uM, data = na.omit(data), fct = G.4()), mean(data$Response), type = "absolute", display = TRUE, multcomp = FALSE)[1]

```

# Figures

```{R}
name = "drug2"
data = subset(Data, Drug == "drug2")
figure = ggplot(data, aes(x = CONCEN.uM, y = Response)) + geom_point()
figure = figure + scale_x_continuous(name = paste("[", capitalize(name), "]", " (uM)"), trans='log') + scale_y_continuous(name = "Response", limits = c(0, max(data$Response)))
figure = figure + theme_bw() + theme_classic()
Con = data.frame(seq(min(data$CONCEN.uM), max(data$CONCEN.uM), length.out = 1000))
Fit1 = data.frame(predict(m1_d1, newdata = Con))
Fit1 = data.frame(Con, Fit1)
names(Fit1) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit1, color = "red")

Fit2 = data.frame(predict(m2_d1, newdata = Con))
Fit2 = data.frame(Con, Fit2)
names(Fit2) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit2, color = "blue")

Fit3 = data.frame(predict(m3_d1, newdata = Con))
Fit3 = data.frame(Con, Fit3)
names(Fit3) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit3, color = "green")

Fit4 = data.frame(predict(m4_d1, newdata = Con))
Fit4 = data.frame(Con, Fit4)
names(Fit4) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit4, color = "purple")

Fit5 = data.frame(predict(m5_d1, newdata = Con))
Fit5 = data.frame(Con, Fit5)
names(Fit5) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit5, color = "cyan")

Fit6 = data.frame(predict(m6_d1, newdata = Con))
Fit6 = data.frame(Con, Fit6)
names(Fit6) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit6, color = "darkgoldenrod")

Fit7 = data.frame(predict(m7_d1, newdata = Con))
Fit7 = data.frame(Con, Fit7)
names(Fit7) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit7, color = "firebrick1")

Fit8 = data.frame(predict(m8_d1, newdata = Con))
Fit8 = data.frame(Con, Fit8)
names(Fit8) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit8, color = "deeppink1")

Fit9 = data.frame(predict(m9_d1, newdata = Con))
Fit9 = data.frame(Con, Fit9)
names(Fit9) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit9, color = "darkorchid4")

Fit10 = data.frame(predict(m10_d1, newdata = Con))
Fit10 = data.frame(Con, Fit10)
names(Fit10) = c("CONCEN.uM", "Response")
figure = figure + geom_line(data = Fit10, color = "darkgreen") 
```

# Smooth spline using ss function in npreg package
## Essentially these are a group of local functions connected by knots 
## The functions could be linear or cubic or other types

```{R}
data$CONCEN.uM1 = data$CONCEN.uM
data$CONCEN.uM1[data$CONCEN.uM == 0] = 0.001
data$logCONCEN.uM = log(data$CONCEN.uM1)

# Linear
mod.sslin = with(data, npreg::ss(logCONCEN.uM, Response, nknots = 9, m =  1))
summary(mod.sslin)

plot(mod.sslin, xlab = "CONCEN.uM", ylab = "Response")
points(data$logCONCEN.uM, data$Response)
xval = approx(x = mod.sslin$y, y = mod.sslin$x, xout = mean(data$Response))$y
exp(xval)


# Cubic
mod.sscub = with(data, npreg::ss(logCONCEN.uM, Response, nknots = 9, m = 2))
summary(mod.sscub)

plot(mod.sscub, xlab = "CONCEN.uM", ylab = "Response")
points(data$logCONCEN.uM, data$Response)
xval = approx(x = mod.sscub$y, y = mod.sscub$x, xout = mean(data$Response))$y
exp(xval)


# quintic
mod.ssqui = with(data, npreg::ss(logCONCEN.uM, Response, nknots = 9, m = 3))
summary(mod.ssqui)

plot(mod.ssqui, xlab = "CONCEN.uM", ylab = "Response")
points(data$logCONCEN.uM, data$Response)
xval = approx(x = mod.ssqui$y, y = mod.ssqui$x, xout = mean(data$Response))$y
exp(xval)
```


********************************************************************************
# Wrap Up
## Show Citations & Version Information
``` {r show_citations}
sessionInfo()
citation("prettyR")
citation("descr")
citation("lubridate")
citation("sjlabelled")
citation("haven")
citation("plyr")
citation("knitr")
citation("car")
citation("dplyr")
citation("npreg")
citation("drc")
```

## Show Timing
``` {r show_timing}
# Show when script started, when it completed, & how long it took to run.
print(Timing <- within(Timing, {End <- Sys.time(); Duration <- End - Start}))

# END OF SCRIPT.
```
